---
title: "S20 STA 100 A05 Discussion 02"
author: "Yishan Huang"
date: "2020/04/07"
output:
  html_document: default
  pdf_document: default
---

Discussion Time: Tuesday 12:10 -- 1:00 pm.

Email Address: yishuang@ucdavis.edu

Zoom: https://ucdstats.zoom.us/j/516752243?pwd=WnhZY3E2SFNxZWhiRHRPcWFrV2hZZz09

Office Hour: Thursday 12:00 -- 1:00 pm.

## Main points in this notes

- Install packages
- Example of dataset, build-in dataset in R
- Dataframe
- Choose rows and columns from a dataframe, subsampling
- Histograms
- Put multiple plots in a single figure
- Add a histogram on an existing plot
- Set range of $x$ and $y$ coordinate in plots
- Main title, labels, legends
- Draw vertical or horizontal lines, or lines with given slopes and intercepts
- Frequency table for 2 categorical variables
- Mosaic plots

## How to install a package

```{r}
# install.packages("ggplot2")
library(ggplot2)
```

## Data

Today we are going to talk about data manipulation process, and deal with some real dataset. We will use one of the most popular build-in dataset 'mtcars' as an example. Let's first have a look on the dataset: 

```{r}
mtcars_data <- mtcars
mtcars_data
# ?mtcars
```

## Subset the data

The type of this data is 'dataframe'. Typically, most of the dataset in R are storaged in the format of dataframe. So it is essential to know fundamental usages of dataframe. For example if we need to take some rows or columns from the whole dataset, we could do the following:

```{r}
mtcars_data[1, ]  # the first row
mtcars_data[, c(2, 3)]  # the second and third column
mtcars_data[, c("cyl", "disp")]
mtcars_data$mpg
mtcars_data[3, 5]  # one specific value
```

Let's do it fancier:

```{r}
mtcars_sub1 <- mtcars_data[mtcars_data$cyl == 6, ]
mtcars_sub2 <- mtcars_data[mtcars_data$cyl == 8, ]
```

## Have a look at the histogram

```{r}
par(mfrow = c(1, 2))
hist1 <- hist(mtcars_sub1$mpg, breaks = 17:22, xlab = "mpg", main = "Histogram of mpg for 6-cylinder cars")
hist2 <- hist(mtcars_sub2$mpg, breaks = 10:20, xlab = "mpg", main = "Histogram of mpg for 8-cylinder cars")

library(scales)
library(RColorBrewer)
pal <- brewer.pal(8, 'Dark2')
par(mfrow = c(1, 1))
plot(hist1, col = alpha(pal[1], 0.5), xlim = c(10, 22), ylim = c(0, 4),
     xlab = "mpg", main = "Histogram of mpg")  # transparency = 0.5
plot(hist2, col = alpha(pal[2], 0.5), add = T)
legend("topright", legend = c("6 cylinders", "8 cylinders"), fill = alpha(pal[1:2], 0.5))

abline(v = c(mean(mtcars_sub1$mpg), mean(mtcars_sub2$mpg)), col = pal[3], lwd = 5)
abline(v = c(median(mtcars_sub1$mpg), median(mtcars_sub2$mpg)), col = pal[4], lty = 2, lwd = 5)

c(mean = mean(mtcars_sub1$mpg), median = median(mtcars_sub1$mpg))
c(mean = mean(mtcars_sub2$mpg), median = median(mtcars_sub2$mpg))
```

## Create mosaic plots

For dataset with two categorical variables, we can create the mosaic plot: 

```{r}
cyl_am_table <- table(mtcars_data[, c("cyl", "am")])
colnames(cyl_am_table) <- c("automatic", "manual")
mosaicplot(cyl_am_table, main = "Mosaic plot for cylindar and transmission", col = pal[1:2])
```



